Investigating the correlation between a vote for Biden or Trump based on a persons education, gender, race and age.

### 2024-03-16 paper/paper.qmd
fixed ages
fixed

raw_data$birthyr <- as.numeric(raw_data$birthyr)
current_year <- as.numeric(year(as.Date(Sys.time())))

@@

# formatting data, mapping race ids to races.
ces2020 <-
  ces2020 |>
  filter(votereg == 1,
         CC20_410 %in% c(1, 2)) |>
  mutate(
    voted_for = if_else(CC20_410 == 1, "Biden", "Trump"),
    voted_for = as_factor(voted_for),
    gender = if_else(gender == 1, "Male", "Female"),
    education = case_when(
      educ == 1 ~ "No HS",
      educ == 2 ~ "High school graduate",
      educ == 3 ~ "Some college",
      educ == 4 ~ "2-year",
      educ == 5 ~ "4-year",
      educ == 6 ~ "Post-grad"
    ),
    education = factor(
      education,
      levels = c(
        "No HS",
        "High school graduate",
        "Some college",
        "2-year",
        "4-year",
        "Post-grad"
      )
    ),
    race = case_when(
      race == 1 ~ "White",
      race == 2 ~ "Black",
      race == 3 ~ "Hispanic",
      race == 4 ~ "Asian",
      race == 5 ~ "Native American",
      race == 6 ~ "Middle Eastern",
      race == 7 ~ "Two or more races",
      race == 8 ~ "Other"
    ),
    race = factor(
      race,
      levels = c(
        "White",
        "Black",
        "Hispanic",
        "Asian",
        "Native American",
        "Middle Eastern",
        "Two or more races",
        "Other"
      )
    )
  ) |> select(voted_for, gender, education, race, age)
cat("Table 4. Header for Processed 2020 US Election Survey data.")
# preview data after mutation.
kable(head(ces2020))
```
TODO: complete bar chart section from the worked example.
# Model
Since we wish to predict how likely it is for a respondent to cast a vote for Biden based on their gender, age, education and race, we shall use the rstanarm (citation needed) library to

@@

```{r}
#| include: true
#| warning: false
#| message: false
#| echo: false
cat("results")
political_preferences <-
  readRDS(file = "~/linear_model_investigation/data/political_preferences.rds")

# model summary
final_table = modelsummary(
  list(
    "Support Biden" = political_preferences
  ),
  statistic = "mad"
)
kable(final_table)
```

>>>>

raw_data$birthyr <- as.numeric(raw_data$birthyr)
current_year <- 2020

@@

# formatting data, mapping race ids to races.
ces2020 <-
  ces2020 |>
  filter(votereg == 1,
         CC20_410 %in% c(1, 2)) |>
  mutate(
    voted_for = if_else(CC20_410 == 1, "Biden", "Trump"),
    voted_for = as_factor(voted_for),
    gender = if_else(gender == 1, "Male", "Female"),
    education = case_when(
      educ == 1 ~ "No HS",
      educ == 2 ~ "High school graduate",
      educ == 3 ~ "Some college",
      educ == 4 ~ "2-year",
      educ == 5 ~ "4-year",
      educ == 6 ~ "Post-grad"
    ),
    education = factor(
      education,
      levels = c(
        "No HS",
        "High school graduate",
        "Some college",
        "2-year",
        "4-year",
        "Post-grad"
      )
    ),
    race = case_when(
      race == 1 ~ "White",
      race == 2 ~ "Black",
      race == 3 ~ "Hispanic",
      race == 4 ~ "Asian",
      race == 5 ~ "Native American",
      race == 6 ~ "Middle Eastern",
      race == 7 ~ "Two or more races",
      race == 8 ~ "Other"
    ),
    race = factor(
      race,
      levels = c(
        "White",
        "Black",
        "Hispanic",
        "Asian",
        "Native American",
        "Middle Eastern",
        "Two or more races",
        "Other"
      )
    )
  ) |> select(voted_for, gender, education, race, age)
cat("Table 4. Header for Processed 2020 US Election Survey data.")
# preview data after mutation.
kable(head(ces2020))
```
TODO: complete bar chart section from the worked example.

# Model
Since we wish to predict how likely it is for a respondent to cast a vote for Biden based on their gender, age, education and race, we shall use the rstanarm (citation needed) library to

@@

```{r}
#| include: true
#| warning: false
#| message: false
#| eval: true
#| echo: false
cat("results")
political_preferences <-
  readRDS(file = "~/linear_model_investigation/data/political_preferences.rds")

# model summary
modelsummary::modelsummary(
  list(
    "Support Biden" = political_preferences
  ),
  statistic = "mad"
)
```
